---
title: "Analyzing R users"
output:
  html_notebook: default
  pdf_document: default
---



## Overview

Stackoverflow recently released the anonymized results of their 2019 annual developer survey. The goal of this project is to analyse various trends for R developers worldwide.



First, let's filter all the R users from the dataset



```{r message=FALSE,warning=FALSE}

#reading necessary libraries
require(tidyverse)
require(countrycode)
require(highcharter)
require(ggplot2)

#reading the data
data <- read_csv("data/survey_results_public.csv")
attach(data)



RUsers<- data %>% 
  filter(grepl('^R$', LanguageWorkedWith) | grepl(';R$', LanguageWorkedWith) | 
           grepl(';R;', LanguageWorkedWith) | grepl('^R;', LanguageWorkedWith)) %>% 
  filter(MainBranch %in% c('I am a developer by profession', 
                           'I am not primarily a developer, but I write code sometimes as part of my work')) %>% 
  filter(Employment %in% c('Employed full-time', 
                           'Employed part-time', 
                           'Independent contractor, freelancer, or self-employed')) %>% 
  filter(!grepl('Other Country', Country)) %>% 
  filter(!is.na(Country), !is.na(ConvertedComp), ConvertedComp > 0) 


attach(RUsers)

```

Let's begin with *Uni-variate* analysis of variables, then slowly we can proceed with *bi-variate* and *multi-variate* analysis as well as soon as things start getting interesting.

### Let's start with the Profession of R Users


```{r message=FALSE,warning=FALSE}

table(MainBranch)


```

So we can notice that most of the R users were developers.


Let's check which country has the most R users


```{r message=FALSE,warning=FALSE}

RCountry <- RUsers %>% group_by(Country) %>% 
  select(Country) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(Count)) %>% 
  top_n(20)

```

We can see that most R users are in US followed by Germany , UK then India.


#### Let's check what do R users professionaly do and their Employment as well: 


```{r message=FALSE, warning=FALSE}

table(MainBranch)
table(Employment)

```
So as we can notice, most of them are __developers__ followed by __students__ and most of them are employed as Full time employees.


### Let's now check the Education level of R users worldwide


```{r warning=FALSE,message=FALSE}

table(EdLevel)

```
So as we can see most of the R users hold  Master's degree followed by Bachelor's degree followed by doctoral degrees as well.

Let's also see the undergraduate degrees of the users.

```{r message=FALSE,warning=FALSE}

table(UndergradMajor)


```
Most of them hold a UG degree in CS and software engineering followed by mathematics and statistics.



### let's analyze educations level and UG for top countries


1) Let's start off with US

```{r warning=FALSE,message=FALSE}

USEduR <- RUsers %>% filter(Country=='United States') %>% 
                    select(EdLevel) %>% 
                    group_by(EdLevel) %>% 
                    summarise(CountEdu=n()) %>% 
                    arrange(desc(CountEdu))


# also doing the same to find the count of UG studies of R users in US
US_UG_R <- RUsers %>% filter(Country=='United States') %>% 
                    select(UndergradMajor) %>% 
                    group_by(UndergradMajor) %>% 
                    summarise(CountUG=n()) %>% 
                    arrange(desc(CountUG))


# Now let's plot them

hchart(USEduR,hcaes(x=USEduR$EdLevel,y=USEduR$CountEdu),type="column",color="purple",name="Count") %>%  
  hc_exporting(enabled = TRUE) %>%
  hc_title(text="Education of R users in US",align="center") %>%
  hc_add_theme(hc_theme_elementary()) 

hchart(US_UG_R,hcaes(x=US_UG_R$UndergradMajor,y=US_UG_R$CountUG),type="column",color="yellow",name="Count") %>%  
  hc_exporting(enabled = TRUE) %>%
  hc_title(text="UG majors of R users in US",align="center") %>%
  hc_add_theme(hc_theme_elementary()) 
 

```


2) Let's check for India now:
```{r warning=FALSE,message=FALSE}
# EdLevel

IndiaEduR <- RUsers %>% filter(Country=='India') %>% 
                    select(EdLevel) %>% 
                    group_by(EdLevel) %>% 
                    summarise(CountEdu=n()) %>% 
                    arrange(desc(CountEdu))


# also doing the same to find the count of UG studies of R users in US
India_UG_R <- RUsers %>% filter(Country=='India') %>% 
                    select(UndergradMajor) %>% 
                    group_by(UndergradMajor) %>% 
                    summarise(CountUG=n()) %>% 
                    arrange(desc(CountUG))


# Now let's plot them

hchart(IndiaEduR,hcaes(x=IndiaEduR$EdLevel,y=IndiaEduR$CountEdu),type="column",color="green",name="Count") %>%  
  hc_exporting(enabled = TRUE) %>%
  hc_title(text="Education of R users in US",align="center") %>%
  hc_add_theme(hc_theme_elementary()) 

hchart(India_UG_R,hcaes(x=India_UG_R$UndergradMajor,y=India_UG_R$CountUG),type="column",color="orange",name="Count") %>%  
  hc_exporting(enabled = TRUE) %>%
  hc_title(text="UG majors of R users in US",align="center") %>%
  hc_add_theme(hc_theme_elementary()) 
 

```


3)Let's check for UK:
```{r warning=FALSE,message=FALSE}
UKEduR <- RUsers %>% filter(Country=='United Kingdom') %>% 
                    select(EdLevel) %>% 
                    group_by(EdLevel) %>% 
                    summarise(CountEdu=n()) %>% 
                    arrange(desc(CountEdu))


# also doing the same to find the count of UG studies of R users in US
UK_UG_R <- RUsers %>% filter(Country=='United Kingdom') %>% 
                    select(UndergradMajor) %>% 
                    group_by(UndergradMajor) %>% 
                    summarise(CountUG=n()) %>% 
                    arrange(desc(CountUG))


# Now let's plot them

hchart(USEduR,hcaes(x=USEduR$EdLevel,y=USEduR$CountEdu),type="column",color="purple",name="Count") %>%  
  hc_exporting(enabled = TRUE) %>%
  hc_title(text="Education of R users in UK",align="center") %>%
  hc_add_theme(hc_theme_elementary()) 

hchart(US_UG_R,hcaes(x=US_UG_R$UndergradMajor,y=US_UG_R$CountUG),type="column",color="yellow",name="Count") %>%  
  hc_exporting(enabled = TRUE) %>%
  hc_title(text="UG majors of R users in UK",align="center") %>%
  hc_add_theme(hc_theme_elementary()) 
 
```


### Let's now analyze the Salaries of R users

Let's start by  making a data table of R users country wise and ther median annual salary arranged by descending order of most R users country wise.

```{r message=FALSE,warning=FALSE}

country_r_users_salary<- RUsers %>%  group_by(Country) %>% 
  summarise(Count=n(), median_salary=round(median(ConvertedComp))) %>% 
  arrange(desc(Count))

```


Let's create another table which is going to contain country and continent information. We will use a __countrycode()__ function to create  a new variable with the name of the continent or region to which each country belongs.

```{r message=FALSE,warning=FALSE}

#mutate() adds new variables while preserving the old ones and distinct() keeps inly the distinct values.
Countries_db <- data %>% filter(!grepl('Other',Country)) %>% 
  distinct(Country) %>% 
  mutate(continent = countrycode(sourcevar = Country, origin = "country.name", destination = "continent"))

```


Let's now plot and analyze the salary distribution country wide.
ConvertedComp is defined as “salary converted to annual USD salaries using the exchange rate on 2019-02-01, assuming 12 working months and 50 working weeks."


Now, let's plot a boxplot of Countries and the median annual salaries of R users, colored by continent.

```{r warning=FALSE,message=FALSE}

#let's modify the R table first
RUsers %>% left_join(Countries_db, by="Country") %>% 
  inner_join(country_r_users_salary %>%  filter(Count >= 10), by = "Country") %>% 
  mutate(Country = reorder(Country,ConvertedComp, fill = continent)) %>% 
  ggplot(aes(x = Country, y = ConvertedComp, fill = continent)) + 
  geom_boxplot(outlier.size = 0.5) +
  ylab('Annual USD salary') +
  coord_flip(ylim = c(0, 200000)) +
  scale_y_continuous(breaks = seq(0, 200000, by = 20000), 
                     labels = function(x) format(x, big.mark = ",", decimal.mark = '.', scientific = FALSE)
                     ) +
  ggtitle("R Users' Salaries by Country") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_discrete(name = "Continent")

  


```
In the above plot the black lines in the boxes represent the median values i.e 50% of the population lies below it and 50% above it.

We can notice that __United States__ offers the highest median salaries to R users of around __110,000 USD__, followed by Israel and Norway. And the lower half is filled by Asian,South American and African countries.



### Let's now check the salary distribution Country wise





#### 1) United States

```{r warning=FALSE,message=FALSE}

RUsers %>%  filter(Country == 'United States') %>% 
            filter(!is.na(Age)) %>% 
            filter(Age >=20) %>% 
            ggplot(aes(x=as.factor(floor(Age/10)), y = ConvertedComp)) + 
            geom_boxplot(outlier.shape = NA) + 
            coord_cartesian(ylim = c(0,300000)) + 
            geom_jitter(aes(as.factor(floor(Age/10)),ConvertedComp), 
                        position= position_jitter(width = 0.1, height = 0),
                        alpha=0.8,
                        size=1) + 
            scale_y_continuous(breaks = seq(0,300000,by=20000),
                               labels = function(x) format(x,big.mark = ",",decimal.mark ='.', scientific = FALSE)) +
            ylab('Annual USD salaries') + 
            xlab('Age bin(years)') + 
            ggtitle("Distributions of R users Salaries by Age bin for US") + 
            theme(plot.title = element_text(hjust = 0.5)) +
            scale_x_discrete(labels=c("2" = "20-29", "3" = "30-39",
                          "4" = "40-49", "5" = "50-59", "6" = "60-69"))
                                 



```
So we can see that the highest salaries is for ages between 50-60 which is expected.



#### 2) India 


```{r warning=FALSE,message=FALSE}


```
