---
title: "Analyzing R user Trends"
output: html_notebook
---

## Overview

Stackoverflow recently released the anonymized results of their 2019 annual developer survey. The goal of this project is to analyse various trends for R users and developers worldwide.



#### Data Wrangling 

The dataset has 90,000 rows. Our goal is to filter R responses so that we can drill down further and analyze the trends related to R developers.

```{r message=FALSE, warning=FALSE}

#reading necessary libraries
require(tidyverse)
require(countrycode)
require(highcharter)
require(ggplot2)

#reading the data
data <- read_csv("data/survey_results_public.csv")
attach(data)

#Let's check who all are developers
devTable <- as.data.frame(table(data$MainBranch))


```

Let's create a braplot of the profession of the respondents.

```{r message=FALSE, warning=FALSE}

hchart(devTable,hcaes(x=Var1,y=Freq),type="column",name="Count",color="#9B6ED8") %>%  
  hc_exporting(enabled = TRUE) %>%
  hc_title(text="Barplot of occupation of participants",align="center") %>%
  hc_add_theme(hc_theme_elementary()) 


```

So we can notice that that most of the participants are <b>developers</b>.


Now let's see how many developers have developement just as a hobby.
```{r message=FALSE, warning=FALSE} 

table(data$Hobbyist)
table(OpenSourcer)

```


Let's check the Employment of the respondents.

```{r message=FALSE, warning=FALSE}

round(prop.table(table(Employment))*100,2)

```
We can notice that 74 % of the respondents were ***Full time employees***.


Now let's check to which country the respondents belonged.

```{r message=FALSE,warning=FALSE}

#Creating a data frame of the Country and its frequency.
countryDf <- data %>% count(Country) %>% 
  top_n(10)

#arranging in descending order of count
countryDf<-arrange(countryDf,desc(n))
  
hchart(countryDf,hcaes(x=Country,y=n),type="column",name="Count") %>%  
  hc_exporting(enabled = TRUE) %>%
  hc_title(text="Barplot of Country of respondents",align="center") %>%
  hc_add_theme(hc_theme_elementary()) 

```
From the above plot we can notice that most of the respondents were from US,India followed by Germany.

```{r message=FALSE,warning=FALSE}
table(Student)

```
We can notice that most of the respondents are not Students.


### Let's see which country had most respondents as Students ?

To achieve this let's create a Dataframe of countries with highest number of Student respondents.

```{r message=FALSE,warning=FALSE}

StudentDf <- data %>% filter(Student=="Yes, full-time") %>% 
  count(Country) %>% 
  top_n(30) %>% 
  arrange(desc(n))

StudentDf


```
F

